Rapid response of mesophyll conductance to light availability allows shade leaves to take advantage of sunflecks

### 2015-11-23 manuscript/manuscript_wtc3.Rmd
comments MT

\
Climate warming may also potentially affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) scale positively with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling A for sun and shade leaves. Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.

@@

\
1.	If whole tree canopies are optimized for C gain, then leaf N, K and photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.

@@

\
3.	As shade leaves are constrained by their underlying biochemistry and slow physiological responses, increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.

\
4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months.

@@

where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite. Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al. [-@evans1986carbon]:

@@

Leaf level instantaneous transpiration efficiency (ITE) was calculated as A divided by E. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near].

##Data analysis
Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4]. For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect.
# Results

@@

# List of Tables
**Table 1**. *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect of leaf type and temperature treatment on each trait.

\
**Table 2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect of leaf type, light environment and temperature treatment on each parameter.

@@

# Tables {.unnumbered}
**Table 1**. *Eucalyptus tereticornis* leaf traits and the distribution of resources between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect of leaf type and temperature treatment on each trait.

@@

\
**Table 2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect of leaf type, light environment and temperature treatment on each parameter.

>>>>

\
Climate warming may also affect the physiological behaviour of leaves within a canopy. This is because leaves can be exposed to different heat, water and high light stresses as temperature and vapour pressure deficit (VPD) vary with canopy light availability [@baldocchi2002environment; @valladares2004photosynthetic; @niinemets2007photosynthesis]. How these stresses affect the diffusion of CO~2~, through either g~s~ or g~m~, will have implications for upscaling A for sun and shade leaves. Additionally, light saturated rates of A are limited by the maximum rate of Rubisco carboxylation (Vc~max~) or the maximum rate of photosynthetic electron transport (J~max~) across a range of temperatures, yet their temperature dependencies are not the same [@farquhar1980biochemical; @medlyn2002temperature]. How these parameters are differentially affected by warming may impact constraints of N distribution and leaf photosynthetic capacity across light gradients. The impacts of warming on plant physiological processes are obviously vast, yet differentiating their impacts on leaf physiology within a canopy will be essential to evaluate whole tree responses to a changing climate.

@@

\
1.	If whole tree canopies are optimized for C gain, then leaf N, hydraulic conductance and biochemical photosynthetic capacity are predicted to be higher in sun leaves compared to shade leaves.

@@

\
3. As shade leaves were expected to develop lower biochemical photosynthetic capacity and leaf physiolgoical responses to light are typically slow (minutes), increases in A following sunfleck simulations were not expected to reach rates of full sun leaves.

\
4.	The effects of climate warming were predicted to be greater in sun than shade leaves, seen as a decrease in g~s~ and leaf C gain during summer months, as increased evaporative demand from higher temperatures will lead to stomatal closure.

@@

where R~s~ is the isotopic ratio of the sample and R~stnd~ is the isotopic ratio of the standard Vienna Pee Dee Belemnite (VPDP). Next, the observed discrimination ($\Delta$~obs~) is calculated from Evans et al. [-@evans1986carbon]:

@@

Leaf level instantaneous transpiration efficiency (ITE) was calculated as A divided by E. The g~1~ parameter was estimated from ITE to VPD response curves by fitting a rearranged optimal g~s~ model for ITE [@medlyn2011reconciling] using non-linear regression, where k=0.5 [see @duursma2013near]. Small values of g~1~ indicate that transpiration is costly in C terms and plants are then likely to exhibit conservative water use [@lin2015optimal].

##Data analysis
Differences in experimental parameters to either the warming treatment or leaf type were analysed by mixed-effects models in R [@RDevelopmentCoreTeam2011] with WTC as a random effect. Explained variance (R^2^) of mixed models were computed as in  [@nakagawa2013general]. Confidence intervals (95 %) of mixed effect linear models were generated using bootstrapping methods with 999 simulations, using the bootMer function in the 'lme4' package [@bates2013lme4]. For non-linear relationships, confidence intervals were estimated by fitting a generalized additive model to the data with the 'mgcv' package, using WTC as a random effect. All tests of statistical significance were conducted at an alpha of 0.05.

# Results

@@

# List of Tables
**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

\
**Table 2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.

@@

# Tables {.unnumbered}
**Table 1**. *Eucalyptus tereticornis* leaf morphological and physiological traits between full sun and shade leaves under ambient and elevated temperature treatments. Leaf mass per area, N~a~, $\delta$^13^C, $\Psi$~pd~, $\Psi$~l~ and K values represent treatment mean (&pm; 1 standard error) across measurement campaigns (n=6). Values of Vc~max~ and J~max~ are treatment mean (&pm; 1 standard error) from AC~i~ curves measured in each chamber at saturating light. Units of LMA and Leaf Nitrogen~area~ are g m^2^, K is mmol m^-2^ s^-1^ MPa, WP is MPA and $\delta$^13^C is &permil;. Different letters represent significant differences between leaf type and temperature treatments. The P value represents the overall effect between each unique combination of leaf type and temperature treatment for each trait.

@@

\
**Table 2**. Responses of *Eucalyptus tereticornis* leaf gas exchange parameters for sun and shade leaves under ambient and elevated temperature treatments. Each value reflects the mean (&pm; 1 standard error) for each treatment across gas exchange campaigns (n=6). Units for A and E are $\mu$mol m^-2^ s^-1^, for g~s~ and g~m~ are mol m^-2^ s^-1^ and for VPD are kPa. Different letters represent significant differences between leaf type, light environment and temperature treatments. The P value represents the overall effect between each unique combination of leaf type, light environment and temperature treatment for each parameter.
